% ============================================================
\section{Computational Methods}
\label{sec:methods}
% ============================================================

\subsection{Pipeline Overview}

The pipeline executes sequentially in 16 numbered steps.
Each step caches its result to disk (JSON or NumPy \texttt{.npz});
on re-run with identical inputs only missing steps are recomputed.
The high-level sequence is:

\begin{enumerate}[nosep]
  \item Download and mosaic Copernicus DEM tiles.
  \item Compute terrain derivatives (slope, aspect, horizon profile,
        TWI — cached).
  \item Download ESA WorldCover and OSM vector features.
  \item Compute \emph{preliminary} $\rzz$ and $\zzz$ at default
        $\thetav = 0.20$ and $\lw = 0$.
  \item Download SoilGrids soil properties for the footprint.
  \item Compute actual $\lw$ from clay fraction (Eq.~\eqref{eq:lw_clay}),
        recompute $\zzz$ and $\rzz$.
  \item Compute $\kT$ (3-D ray-casting over DEM).
  \item Compute $\kM$ (horizon-based muon integral).
  \item Compute $\kL$ (WorldCover + OSM, $W(r)$-weighted).
  \item Compute site neutron flux and $N_0$
        (Eqs.~\eqref{eq:Nneut_site}--\eqref{eq:N0}).
  \item Download ERA5 climatology and compute
        $\rho_{WV}$, $f_{WV}$, SWE (monthly).
  \item Download MODIS LAI and compute AGBH.
  \item Query Macrostrat for geological characterisation.
  \item Download ERA5-Land hourly soil moisture (year by year,
        cached); compute multi-year monthly statistics.
  \item Downscale and fuse ERA5 soil moisture with local pedology,
        topography, and land cover; propagate uncertainty.
  \item Compute optimal soil sampling plan and $\rzz$ dynamic range.
  \item Assemble the full report (PDF + JSON).
\end{enumerate}

\subsection{Terrain Processing}
\label{sec:methods:terrain}

\paragraph{Slope and aspect.}
Slope $\beta$ and aspect are derived from the DEM by the standard
central-difference gradient:
\begin{equation}
  \nabla z = \left(\frac{\partial z}{\partial x},\,
                   \frac{\partial z}{\partial y}\right),
  \qquad
  \beta = \arctan|\nabla z|.
\end{equation}
A minimum slope floor of $\SI{0.01}{\degree}$ is applied to avoid
division by zero in subsequent hydrological calculations.

\paragraph{Horizon profile.}
For each of $N_\phi = 72$ azimuth directions (5° spacing), the
maximum elevation angle $\psi(\phi)$ is computed by scanning along
the DEM at that azimuth out to $3\,\rzz$ from the site.
The result is a 72-element array $\psi(\phi_k)$ that enters both
the $\kM$ integral and the ray-casting algorithm.

\paragraph{Topographic Wetness Index.}
TWI is computed as \citep{beven1979}:
\begin{equation}
  \mathrm{TWI} = \ln\frac{A_s}{\tan\beta},
  \label{eq:twi}
\end{equation}
where $A_s$ is the specific upslope drainage area per unit contour
length, derived from the $D_\infty$ flow accumulation of the
pit-filled DEM \citep{wang2006}.
Because TWI computation is the most expensive terrain step
($\mathcal{O}(N_\mathrm{pix}\log N_\mathrm{pix})$ for large DEMs),
results are cached to disk as a \texttt{.npz} file.
The cache key is a hash of the DEM shape, pixel spacing, $\rzz$,
and summary statistics (mean and standard deviation of elevation);
if the hash matches the stored file the cached result is returned
directly.

\subsection{Topographic Correction \texorpdfstring{$\kT$}{kappa\_topo}}
\label{sec:methods:kT}

The 3-D ray-casting implementation samples $N_\theta = 18$ zenith
angles (uniformly in $\cos\theta$, $0^\circ$–$89^\circ$) and
$N_\phi = 72$ azimuth angles.
For each ray $(\theta_k, \phi_j)$:
\begin{enumerate}[nosep]
  \item March along the ray in 3-D until it intersects the DEM
        surface, recording the air path length $L_\mathrm{air}$.
  \item At the intersection, read the surface slope $\alpha$ and
        compute $L_\mathrm{soil} = \zzz / \cos\alpha$.
  \item Evaluate Eq.~\eqref{eq:ray}.
\end{enumerate}
The flat-reference integral uses a perfectly horizontal DEM at
the sensor elevation.
Both integrals are computed with the same angular sampling;
$\kT$ is their ratio.

Rays that exit the DEM bounding box without hitting terrain
(i.e., the site is on a ridge or cliff edge) contribute zero to the
numerator but retain their full flat-reference contribution in the
denominator, correctly reducing $\kT$ below 1.

\subsection{LULC Correction and Supplementary Hydrogen}
\label{sec:methods:kL}

\paragraph{WorldCover raster.}
Each 10-m WorldCover pixel within $2\,\rzz$ of the site is assigned
the hydrogen equivalence factor $f_H^{(c)}$ from Table~\ref{tab:fH}.
The radial weight $W(r_i)$ (Eq.~\eqref{eq:Wr}) is evaluated at the
pixel centroid distance $r_i$.
The correction factor $\kL$ follows Eq.~\eqref{eq:kappa_lulc}.

\paragraph{OSM vector features.}
OSM polygons (buildings, water bodies, sealed roads) are rasterised
onto the DEM grid at 1-m resolution and then downsampled to the
DEM pixel size.
For pixels with OSM overlap the WorldCover $f_H$ is replaced by the
OSM-specific value (e.g.\ $f_H = 4.0$ for permanent water,
$f_H = 0.05$ for sealed surfaces).
The uncovered area inside the $2\,\rzz$ circle contributes
\begin{equation}
  \mathrm{base\_contrib} =
    f_H^\mathrm{base} \times
    \max\!\bigl(0,\;
      \textstyle\sum_i W(r_i)\,A_i - \sum_{j\in\mathrm{OSM}} W(r_j)\,A_j
    \bigr),
  \label{eq:osm_baseline}
\end{equation}
where $f_H^\mathrm{base}$ is the WorldCover bare-soil baseline
($f_H = 0.03$) and the difference of $W$-weighted integrals
avoids the spatially averaged approximation used in simpler
implementations.

\paragraph{Reference moisture.}
All $f_H$ values and the WorldCover baseline are computed at the
user-specified initial soil moisture $\thetav^\mathrm{ref}$
(\texttt{THETA\_V\_INIT} in the configuration file, default
$\SI{0.20}{\cubic\metre\per\cubic\metre}$).
This makes $\kL$ physically consistent with the $\rzz$ and
$\zzz$ computed at the same moisture level.

\subsection{Supplementary Correction Factors}
\label{sec:methods:suppl}

These factors are not included in $\kTot$ but are reported separately
as operational correction guidance.

\paragraph{Atmospheric water vapour.}
\label{sec:methods:wv}
An increase in above-sensor water vapour reduces the epithermal
neutron flux reaching the ground.
Following \citet{rosolem2013}, the correction factor is:
\begin{equation}
  f_{WV} = \exp\!\bigl(-\alpha_{WV}\,\rho_{WV}\bigr),
  \label{eq:fWV}
\end{equation}
with $\alpha_{WV} = \SI{0.0054}{\metre\cubed\per\gram}$
\citep{zreda2012}.
Monthly climatological values of $\rho_{WV}$
(Eqs.~\eqref{eq:esat}–\eqref{eq:rhoWV}) and the corresponding
$f_{WV}$ are tabulated in the report.
Note that $f_{WV}$ is an operational (epoch-by-epoch) correction;
the climatology here serves as site-characterisation guidance.

\paragraph{Above-ground biomass hydrogen (AGBH).}
\label{sec:methods:agbh}
Vegetation intercepts and stores water; the associated hydrogen
moderates neutrons \citep{baatz2015, schron2017}.
An effective soil-moisture equivalent is estimated as:
\begin{equation}
  \mathrm{AGBH} = \mathrm{LAI} \times \SI{150}{\gram\per\metre\squared}
                  + \theta_\mathrm{litter}
  \quad[\si{\milli\metre}],
  \label{eq:agbh}
\end{equation}
where the 150 g m$^{-2}$ factor is the mean interception capacity
per unit LAI \citep{schron2017} and $\theta_\mathrm{litter}$ is an
optional litter-water depth.

\paragraph{Snow-water equivalent.}
Snow on the ground effectively acts as liquid water from the
neutron probe's perspective.
Monthly mean SWE (Eq.~\eqref{eq:swe}) is reported alongside
the month in which snow cover first exceeds
$\SI{150}{\milli\metre}$ SWE — a threshold at which the detector
is largely opaque to cosmic-ray neutrons and calibration is
impractical \citep{bogena2017}.

\paragraph{Soil organic carbon hydrogen equivalent.}
Soil organic matter (SOM) contains hydrogen at a concentration
similar to liquid water.
The volumetric hydrogen equivalent is approximated as
\citep{bogena2017}:
\begin{equation}
  \theta_\mathrm{SOC} \approx
    0.55\,\rho_b\,\omega_\mathrm{SOM}
  \quad[\si{\cubic\metre\per\cubic\metre}],
  \label{eq:thetaSOC}
\end{equation}
where $\omega_\mathrm{SOM}$ is the SOM mass fraction [g/g]
(SOM $\approx 1.724\times$ SOC by convention).
This term is reported alongside $\lw$ as a supplementary offset
that does not change with time.

\subsection{Optimal Soil Sampling Plan}
\label{sec:methods:sampling}

A CRNS calibration requires in-situ gravimetric soil moisture
measurements distributed across the footprint in proportion to
the neutron sensitivity kernel $W(r)$.
The pipeline designs a plan of $n_\mathrm{tot} = 26$ sampling
locations arranged in four concentric rings at radii
$r_k = \{0.25,\, 0.50,\, 0.75,\, 1.00\} \times \rzz$.

\paragraph{Ring weight.}
The contribution of ring $k$ to the footprint integral is
\begin{equation}
  \Phi_k = \int_{r_{k-1}}^{r_k} W(r)\,2\pi r\,\mathrm{d}r,
  \label{eq:ring_weight}
\end{equation}
evaluated numerically by the trapezoidal rule.
Points per ring $n_k \in \{4, 6, 8, 8\}$ are allocated so that
$n_k / \sum n_j \approx \Phi_k / \sum \Phi_j$.

\paragraph{Sample depth.}
At each location, soil is sampled at three depth intervals
(0–5, 5–15, 15–30 cm) to resolve the vertical weighting of
neutron penetration.

\paragraph{Footprint dynamic range.}
As supplementary information, the report includes a plot of
$\rzz(\thetav)$ computed over the full moisture range
$[0.01,\, 0.50]$ \si{\cubic\metre\per\cubic\metre}.
This shows how the footprint radius, and hence the spatial
representativeness of the measurements, changes between dry and
wet conditions at the site.

\subsection{Soil Moisture Downscaling and Data Fusion}
\label{sec:methods:smfusion}

ERA5-Land soil moisture has a native resolution of $\sim$9\,km and is
produced with a globally uniform land-surface model.
Local conditions (topography, soil texture, land cover) can shift
the true soil moisture at the CRNS site significantly relative to the
ERA5 grid-cell value.
The module \texttt{smphysics.py} applies a sequential chain of
physically motivated corrections implemented in \texttt{fuse\_soil\_moisture()}.

\paragraph{Step 1 — ERA5 raw signal.}
The 0--7\,cm monthly mean from \texttt{era5sm.py} is used as the
starting estimate, as this layer most closely matches the effective
sensing depth of CRNS at low to moderate moisture.

\paragraph{Step 2 — Pedological rescaling (Saxton \& Rawls 2006).}
ERA5-Land uses globally uniform hydraulic parameters.
Site-specific Field Capacity (FC) and Wilting Point (WP) are derived
from the SoilGrids sand, clay, and organic-matter fractions via the
pedotransfer functions of \citet{saxton2006}:
\begin{equation}
  \theta_\mathrm{local} = \mathrm{WP}_\mathrm{loc}
    + \frac{\theta_\mathrm{ERA5} - \mathrm{WP}_\mathrm{ERA5}}
           {\mathrm{FC}_\mathrm{ERA5} - \mathrm{WP}_\mathrm{ERA5}}
    \left(\mathrm{FC}_\mathrm{loc} - \mathrm{WP}_\mathrm{loc}\right),
  \label{eq:pedol_rescaling}
\end{equation}
with $\mathrm{FC}_\mathrm{ERA5} = 0.32$ and
$\mathrm{WP}_\mathrm{ERA5} = 0.12$\,\si{\cubic\metre\per\cubic\metre}
as representative global defaults.
The result is clipped to $[\mathrm{WP}_\mathrm{loc},\,\mathrm{FC}_\mathrm{loc}]$.

\paragraph{Step 3 — Topographic correction.}
Two terrain effects are applied:
\begin{enumerate}[nosep]
  \item \emph{TWI term (additive):}
        sites with TWI above the footprint mean accumulate more lateral
        flow and are therefore wetter:
        \begin{equation}
          \Delta\theta_\mathrm{TWI}
            = \beta_\mathrm{TWI}
              \left(\mathrm{TWI}_\mathrm{site} - \overline{\mathrm{TWI}}_\mathrm{fp}\right),
          \label{eq:delta_twi}
        \end{equation}
        with default $\beta_\mathrm{TWI} = 0.008$\,\si{\cubic\metre\per\cubic\metre}
        per TWI unit, clipped to $\pm0.15$.
  \item \emph{PISR term (multiplicative):}
        higher annual solar irradiation increases evapotranspiration
        and dries the soil:
        \begin{equation}
          f_\mathrm{PISR}
            = \exp\!\left(-\alpha_\mathrm{PISR}
              \frac{\mathrm{PISR}_\mathrm{site} - \mathrm{PISR}_\mathrm{ERA5}}
                   {\mathrm{PISR}_\mathrm{ERA5}}\right),
          \label{eq:f_pisr}
        \end{equation}
        with default $\alpha_\mathrm{PISR} = 0.12$.
\end{enumerate}
The combined topographic correction is
$\theta_\mathrm{topo} = (\theta_\mathrm{local} + \Delta\theta_\mathrm{TWI})
 \times f_\mathrm{PISR}$.

\paragraph{Step 4 — LULC correction.}
A categorical correction is applied based on WorldCover class fractions
and the $\kappa_\mathrm{LULC}$ already computed for the neutron sensitivity
model:
\begin{itemize}[nosep]
  \item Snow/ice $> 30\%$: $\theta \times 0.4$ (snow is not liquid water);
  \item Impervious (built-up + bare) $> 50\%$: $\theta \times 0.30$;
  \item Impervious $20$--$50\%$: $\theta \times (1 - 0.5\,f_\mathrm{imp})$;
  \item Wetland + forest $> 40\%$ and $\kappa_\mathrm{LULC} > 1.2$:
        $\theta \times 1.10$ (structurally wet surface).
\end{itemize}

\paragraph{Uncertainty.}
A total uncertainty is propagated in quadrature from four independent
components (Gaussian error propagation):
\begin{equation}
  \sigma_\mathrm{total}
    = \sqrt{\sigma_\mathrm{ERA5}^2
          + \sigma_\mathrm{topo}^2
          + \sigma_\mathrm{pedol}^2
          + \sigma_\mathrm{LULC}^2}
    \approx 0.066\;\si{\cubic\metre\per\cubic\metre},
  \label{eq:sigma_sm}
\end{equation}
with default components
$\sigma_\mathrm{ERA5} = 0.05$,
$\sigma_\mathrm{topo} = 0.02$,
$\sigma_\mathrm{pedol} = 0.03$,
$\sigma_\mathrm{LULC} = 0.02$.
The report provides monthly $\theta_v \pm \sigma$ bounds alongside the
ERA5 raw signal.

\subsection{Geological Characterisation}

The Macrostrat API v2 is queried with the site coordinates at three
resolution scales (large, medium, small) in descending order of
coverage.
The first non-empty response is used.
The pipeline extracts:
\begin{itemize}[nosep]
  \item dominant lithology name and Macrostrat rock type
        (carbonate, siliciclastic, igneous, \ldots);
  \item estimated influence on $\lw$ (low, medium, high), based on
        the known hydrogen content of the rock type;
  \item an indicative radiogenicity level (for U/Th-based radiation
        background assessment);
  \item geologic age era (Cenozoic, Mesozoic, Palaeozoic, Precambrian).
\end{itemize}
This qualitative summary complements the quantitative SoilGrids
$\lw$ estimate when geological context is relevant (e.g.\
evaporite-rich or serpentinised terrain).
